Nail: A practical interface generator for data formats

### 2014-11-25 paper/login/design.tex
Login

We will use DNS packets as a leading example, as defined in RFC 1035~\cite[\S4]{RFC:1035}.  Each DNS packet consists of a
header, a set of question records, and a set of answer records. Domain
names in both queries and answers are encoded as a sequence of labels,
terminated by a zero byte.  Labels are Pascal-style strings, consisting
of a length field followed by that many bytes comprising the label.
\label{s:design}
\subsection{Basic Data formats}

\input{fig-dnsfull}
Let's stepping through a simplified Nail grammar for non-compressed
packets, shown in  Figure~\ref{fig:dns-full}.  For this grammar, Nail produces the type declarations
shown in Figure~\ref{fig:dns-full-struct}, and the parser and
generator functions shown in Figure~\ref{fig:dns-full-api}.
Nail grammars are reuseable between applications and we will use this grammar to implement
both a DNS server and client, which previously would have had two separate hand-written parsers,
leading to bugs such as  the Android Master Key.

@@

Just like C supports various constructs to build up types, such as structures, unions, pointers and
elemental types, Nail supports a integers as elemental parsers and a plethora of \emph{combinators}
to represent features of a file or protocol. We will present the features we used in implementing
DNS. A more complete reference can be found in~\ref{fig:syntax}, with detailed rationale in~\cite{bangert:nail-osdi14}.

@@

The body of the \texttt{dnspacket} rule is a structure, which contains any number of fields enclosed
between  curly braces. Each field in the structure is parsed in sequence, and represented as a
structure to the programmer.  Contrary to other programming languages, Nail does not have a special
keyword for structs. We also reverse the usual structure-field syntax: \texttt{id uint1} is a field
called \texttt{id} with type \texttt{uint1}. Often, Nail grammars have
structures with just one non-constant field, for example when parsing a fixed header. Nail supports
this with an alternative form of the structure combinator,
using angle brackets, that contains one unnamed, non-constant field, which is represented directly
in the datatype, without introducing another layer of indirection, as shown on line 40.

@@

Nail provides only one built-in combinator that uses dependent fields, \cc{n_of}, which acts like
the \cc{many} combinator, except it parses an exact number of repetitions specified in the dependent
field. Lines 16-19 in Figure~\ref{fig:dns-full} shows how to use \cc{n_of} to parse the question and
answer records in a DNS packet.
Other dependencies, such as offset fields or checksums, are not handled directly by combinators, but
through  transformations, as we describe next.

\subsection{Input streams and transformations}
\label{s:transforms}
Formats have

Traditional parsers handle input one symbol at a time, from beginning to end.
However, real-world formats often require non-linear parsing. Offset fields require a parser to move
to a different position in the input, possibly backwards. Size fields require the parser to stop
processing before the end of input has been reached, and perhaps resume executing a parent parser.
Other cases, such as compressed data, require more complicated processing on parts of the input
before it can be handled.
Streams represent a sequence of bytes that contain some external format. The parsers and generators
that Nail generates always operate on an implicit stream named \cc{\$current} that they process front to
back, reading input or appending output.
Grammars can use the \cc{apply} combinator to parse a stream or generate output into it. 
Streams are passed as arguments to a rule or defined within the grammar through \emph{transformations}.
The current stream is always passed as an implicit parameter.

There are many ways conceptual streams could be encoded in a data format --
A tool like Nail needs to support all of them. This makes data description languages like
PADS~\cite{Fisher:2005:PDL:1064978.1065046} contain not just a kitchen sink, but a kitchen store
full of features -- and a language that can handle all possible format oddities will be a general
purpose language.  Nail instead of allows you to plug in your own code as \emph{transformations},
with a small, but powerful camping kitchen included that will handle most formats.

Transformations are two arbitrary functions called during parsing and output generation.
The parsing function  takes any number of stream arguments and dependent field values,
and produces any number of temporary streams. This function may reposition and read from the
input streams and read the values of dependent fields, but not change their contents and values. 
The generating function has to be an inverse of the parsing function. It takes the same number of
temporary streams that the parsing function produces, and writes the same number of streams and
dependent field values that the parsing function consumes.

Typically, grammars start out with a single stream, which may then
be broken up by various transformations and passed to sub-rules, which eventually parse various linear
fragment streams. Upon parsing, these fragment streams are generated and then combined by the
transforms.

 To make your life easier and your code safer, Nails provides implementations of
transformations for many common features, such as checksums, size, and offset fields. Furthermore,
Nail provides library functions that can be used to safely operate on streams, such as splitting and
concatenation. Nail implements streams as iterators, so they can share underlying buffers and can be
efficiently duplicated and split.

@@

Transformations can handle a wide variety of patterns in data formats, including the following:

\paragraph{Compressed data.}
For example, DNS uses label compression to reduce the overhead of including each domain name
multiple times in a DNS reply. If a domain name suffix is repeated, it may be replaced by a two-bit
marker followed by a  14-bit offset into the packet, indicating the position where that suffix was
previously encoded.  We use a transform in line 35
of our DNS grammar to handle DNS label compression; the signatures of the two transform functions are shown
in Figure~\ref{fig:dns-xform}.  When parsing a packet, this transform
decompresses the DNS label stream.
When generating a packet, this transform receives the current suffix as
an input, and scans the packet so far for previous occurrences and inserts the offset.

\begin{figure}
\smaller[0.5]
\input{code/dns-xform}
\caption{Signatures of stream transform functions for handling DNS label compression.}
\label{fig:dns-xform}
\end{figure}
Handling label compression in existing tools, such as Bison or Hammer,
would be awkward at best, because some ad-hoc trick would have to
be used to re-position the parser's input stream.  Keeping track of the
position of all recognized labels would not be enough, although still awkward, as the offset field
may refer to any byte within the packet.
For this reason, the sample DNS grammar in Hammer does not
support compression.

\paragraph{Offsets.}
We provide a  transformation for handling offset fields, which is invoked as follows:
\cc{\$fragment transform offset\_u32(\$current, \cc{@}offset)}. 
This transformation corresponds to two functions for parsing and
generation, as shown in Figure~\ref{fig:xform-sig}. It defines a new stream \cc{\$fragment} that can
be used to parse data at the offset contained in \cc{@offset}, by
using \cc{apply \$fragment some_parser}.

\begin{figure}[h]
\smaller[0.5]
\input{code/offset-xform}
\caption{Pseudocode for two functions that implement the offset transform.}
\label{fig:xform-sig}
\end{figure}

\paragraph{Sizes.}
A similar transformation handles size fields. Just like the offset transform, it takes two parameters, a
stream and a dependent field, but instead of returning the suffix of the current stream after an
offset, it returns a slice of the given size from the current stream starting at its current
position. When generating, it appends the fragment stream to the current stream and writes the size
of the fragment to the dependent field.

\paragraph{Checksums.}
Checksums can be verified and computed in a transformation that takes a stream and a dependent
field. In some cases, a checksum is calculated over a buffer that contains the checksum itself, with
the checksum being set to some particular value.

\paragraph{}
A real-world example with many different transforms, used to support
the ZIP file format, is described in our full paper.%\S\ref{s:eval-formats}.

>>>>

We will use DNS packets as a leading example, as defined in RFC 1035.  Each DNS packet consists of a
header, a set of question records, and a set of answer records. Domain
names in both queries and answers are encoded as a sequence of labels,
terminated by a zero byte.  Labels are Pascal-style strings, consisting
of a length field followed by that many bytes comprising the label.
\label{s:design}
\subsection{Basic Data formats}

\input{fig-dnsfull}
Let's step through a simplified Nail grammar DNS
packets, shown in  Figure~\ref{fig:dns-full}.  For this grammar, Nail produces the type declarations
shown in Figure~\ref{fig:dns-full-struct}, and the parser and
generator functions shown in Figure~\ref{fig:dns-full-api}.
Nail grammars are reuseable between applications and we will use this grammar to implement
both a DNS server and client, which previously would have had two separate hand-written parsers,
leading to bugs such as  the Android Master Key.

@@

Just like C supports various constructs to build up types, such as structures and unions from pointers and
elemental types, Nail supports various \emph{combinators} to represent features of a file or protocol. We will present the features we used in implementing
DNS. A more complete reference can be found in~\ref{fig:syntax}, with detailed rationale in~\cite{bangert:nail-osdi14}.

@@

The body of the \texttt{dnspacket} rule is a structure, which contains any number of fields enclosed
between  curly braces. Each field in the structure is parsed in sequence, and represented as a
structure to the programmer.  Contrary to other programming languages, Nail does not have a special
keyword for structs. We also reverse the usual structure-field syntax: \texttt{id uint1} is a field
called \texttt{id} with type \texttt{uint1}. Often, Nail grammars have
structures with just one non-constant field, for example when parsing a fixed header. Nail supports
this with an alternative form of structures,
using angle brackets, that contains one unnamed, non-constant field, which is represented directly
in the datatype, without introducing another layer of indirection, as shown on line 40.

\paragraph{Arrays.}

Nail supports various forms of arrays. Line 40 shows how to parse a domain in a DNS packet with
\cc{many}, which keeps repeating the \cc{label} rule until it fails.  In the next section, we will
explain how to handle count fields, and our full paper describes how to handle various array
representations (such as delimiters or non-empty arrays).

@@

Nail allows handling count fields with \cc{n\_of}, which parses an exact number of repetitions of a rule. Lines 16-19 in Figure~\ref{fig:dns-full} shows how to use \cc{n_of} to parse the question and
answer records in a DNS packet.
Other dependencies, such as offset fields or checksums, are not handled directly by combinators, but
through  transformations, as we describe next.

\subsection{Input streams and transformations}
\label{s:transforms}
So far, we have described a parser that consumes input a byte at a time from beginning to end.
However, real-world formats often require non-linear parsing. Offset fields require a parser to move
to a different position in the input, possibly backwards. Size fields require the parser to stop
processing before the end of input has been reached.
Other cases, such as compressed data and checksums, require more complicated processing on parts of the input
before it can be handled.

For a parser to be useful, it needs to support all these ways of structuring a format. This makes data description languages like
PADS~\cite{Fisher:2005:PDL:1064978.1065046} contain not just a kitchen sink, but a kitchen store
full of features -- and a language that can handle all possible formats will be a general
purpose programming language.  Instead, we keep Nail itself small, and introduce an
interface that allows complicated format structures to be handled by  plug-in \emph{transformations} 
in a general purpose
language. Of course, we ship Nail with a handy library of common transformations to handle common
format features, such as offsets,sizes and checksums.

These \emph{transformations} consume and produce
\emph{streams} -- sequences of bytes -- which can be further passed to other transformations and
eventually parsed by a Nail rule. Transformations can also access values in dependent fields. 
Streams can be subsets of other streams, for example the substream starting at an
offset given in a dependent field to handle pointer fields, or computed at runtime
, such as by decompressing another stream with zlib.

Transformations are two arbitrary functions called during parsing and output generation.
The parsing function consumes any number of streams and dependent field values,
and produces any number of temporary streams. This function may reposition and read from the
input streams and read the values of dependent fields, but not change their contents and values. 
The generating function has to be an inverse of the parsing function, consuming streams and
producing dependent field values and other streams.

\input{fig-zip}

As a concrete example, we will show a grammar for ProtoZIP, a very simple archive format inspired by ZIP in
Figure~\ref{fig:zip-extract}. ProtoZIP consists of a variable-length end-of-file directory, which
is a magic number followed by an array of filenames and pointers to compressed files. A grammar for the real ZIP format, which
has more layers of indirection, is presented in the full paper.

 The grammar first calls the zipdir transform on line 2, which finds the magic number and splits the file into
 two streams, one containing the compressed files, the other the directory.  Streams are referred to
 with \cc{\$identifiers}, similar to dependent fields. A C prototype of the \cc{zipdir} transform is
 shown in\ref{fig:zip-transform}. 
\begin{figure}[tb]
\smaller[0.5]
\input{code/zip-eod-xform}
\caption{Signatures of stream transform functions for handling the
end-to-beginning structure of ProtoZIP files.}
\label{fig:zip-transform}
\end{figure}
When parsing input, this will call \cc{zipdir\_parse}, which takes \cc{\$current} -- an implicit
identifier always refering to the stream currently being handled -- and returns \cc{\$files} and
\cc{\$header}. When generating output, this will call \cc{zipdir\_generate}, which appends
\cc{\$files} and \cc{\$header} to \cc{\$current}.

Line 3 then applies the \cc{dir} rule to  the \cc{\$header} stream, passing it the \cc{\$files}
stream. Within \cc{dir}, \cc{\$current} is now \cc{\$header} and input is parsed from and output
generated to that stream. The \cc{dir} rule in turn describes the structure of the directory -- a
magic number and a count field, followed by that many file descripors. Each file descriptor, is then parsed with two
transformations, the standard-library \cc{slice} which describes an offset and a size within another
stream, and the custom \cc{zlib}, which compresses a stream using zlib. Finally, we apply a trivial
grammar (line 14) to the contents.

In a more complicated example, such as an Office document, we could now specify grammars for each
entry within an archive.

@@

### 2014-11-25 paper/login/eval.tex
Login

\paragraph{DNS} In Section~\ref{s:design}, we introduced Nail's syntax with a grammar
for DNS packets, shown in Figure~\ref{fig:dns-full}.
The grammar corresponds almost directly to the diagrams in RFC 1035,
which defines DNS~\cite[\S4]{RFC:1035}. Nail's dependent fields handle DNS's count fields, and
transformations represent label compression. Both of these features are awkward at best to handle
with existing tools.

\paragraph{ZIP} An especially tricky data format is the ZIP compressed archive format~\cite{pkzip}.
ZIP files are normally parsed end-to-beginning. At the end of each ZIP file is an
\emph{end-of-directory header}. This header contains a variable-length comment, so it has to be
located by scanning backwards from the end of the file until a magic number and a valid length field
is found. Many ZIP implementations disagree on the exact semantics of this, such as when the comment
contains the magic number~\cite{wolf:berlinsides-zip}.

@@

 However, if attackers craft inconsistent files, hand written parsers fail in interesting ways, e.g. 
the length in the central directory is used to allocate memory and the
length in the local directory is used to extract without checking that
they are equal first, as was the case in the Python ZIP
library~\cite{python-bug:20078}.

@@

We ran the DNS fuzzer provided with the Metasploit
framework~\cite{mspframework} on our DNS server, which sent randomly
corrupted DNS queries to our server for 4 hours, during which it did
not crash or trigger the stack or heap corruption detector.

To evaluate whether Nail-based parsers are compatible with good
performance, we compare the performance of our DNS server to that of ISC
BIND 9 release 9.9.5~\cite{bind9}, a mature and widely used DNS server.
We simulate a load resembling that of an authoritative name server, generating a random zone file
and a random sequence of queries, with 10\% non-existent domains. We repeated this sequence of
queries for one minute against both DNS servers. We found that our DNS server is approximately three
times faster than BIND. Although BIND is a more sophisticated DNS server,
and implements many features that are not present in our Nail-based DNS
server and that allow it to be used in more complicated configurations, 
we believe our results demonstrate that Nail's parsers are not
a barrier to achieving good performance.
\paragraph{ZIP.}

>>>>

\paragraph{DNS} In Section~\ref{s:design}, we introduced Nail's syntax with a grammar
for DNS packets, shown in Figure~\ref{fig:dns-full}.
The grammar corresponds almost directly to the diagrams in RFC 1035,
which defines DNS. Nail's dependent fields handle DNS's count fields, and
transformations represent label compression. Both of these features are awkward at best to handle
with existing tools.

\paragraph{ZIP} An especially tricky data format is the ZIP compressed archive format, as specified
by PKWare.
ZIP files are normally parsed end-to-beginning. At the end of each ZIP file is an
\emph{end-of-directory header}. This header contains a variable-length comment, so it has to be
located by scanning backwards from the end of the file until a magic number and a valid length field
is found. Many ZIP implementations disagree on the exact semantics of this, such as when the comment
contains the magic number~\cite{wolf:berlinsides-zip}.

@@

 However, if attackers craft inconsistent files, hand written parsers fail in interesting ways, e.g. 
the length in the central directory is used to allocate memory and the
length in the local directory is used to extract without checking that
they are equal first, as occured in Python bug 20078.

@@

We ran the DNS fuzzer provided with the Metasploit
framework on our DNS server, which sent randomly
corrupted DNS queries to our server for 4 hours, during which it did
not crash or trigger the stack or heap corruption detector.

To evaluate whether Nail-based parsers are compatible with good
performance, we compare the performance of our DNS server to that of ISC
BIND 9 release 9.9.5, a mature and widely used DNS server.
We simulate a load resembling that of an authoritative name server, generating a random zone file
and a random sequence of queries, with 10\% non-existent domains. We repeated this sequence of
queries for one minute against both DNS servers. We found that our DNS server is approximately three
times faster than BIND. Although BIND is a more sophisticated DNS server,
and implements many features that are not present in our Nail-based DNS
server and that allow it to be used in more complicated configurations, 
we believe our results demonstrate that Nail's parsers are not
a barrier to achieving good performance.
\paragraph{ZIP.}

### 2014-11-25 paper/login/intro.tex
Login

  -- Edward Shelley on the Android Master Key
\end{quote}
\section{Introduction}
Binary file formats and network protocols are hard to parse safely: 
The libpng image decompression library had 24 remotely
exploitable vulnerabilities from 2007 to 2013~\cite{cvedetails:libpng},
Adobe's PDF and Flash viewers have been notoriously plagued by input
processing vulnerabilities, and even the zlib compression library had
input processing vulnerabilities in the past~\cite{cvedetails:zlib}.
Most of these attacks involve memory corruption - therefore, it is easy to assume that 
solving memory corruption will end all our woes when handling untrusted inputs.

@@

Furthermore, some parser generators are cumbersome to use when parsing binary data for several
reasons. First, many parser generators don't directly data structures, but call semantic actions
that the you write to construct a data structure. Therefore,you must
describe the format up to three times -- in the grammar, the data structure and the semantic
actions. Second, most parser generators only address parsing inputs, so you'd have to manually
construct outputs. Some parser generators, such as 
Boost.Spirit~\cite{boost-spirit} allow generating output, but require you to write another set of semantic actions.

>>>>

  -- Edward Shelley on the Android Master Key
\end{quote}
\section{Introduction}
Binary file formats and network protocols are hard to parse safely: 
The libpng image decompression library had 24 remotely
exploitable vulnerabilities from 2007 to 2013 according to CVEDetails,
Adobe's PDF and Flash viewers have been notoriously plagued by input
processing vulnerabilities, and even the zlib compression library had
input processing vulnerabilities in the past.
Most of these attacks involve memory corruption - therefore, it is easy to assume that 
solving memory corruption will end all our woes when handling untrusted inputs.

@@

Furthermore, some parser generators are cumbersome to use when parsing binary data for several
reasons. First, many parser generators don't directly data structures, but call semantic actions
that the you write to construct a data structure. Therefore,you must
describe the format up to three times -- in the grammar, the data structure and the semantic
actions. Second, most parser generators only address parsing inputs, so you'd have to manually
construct outputs. Some parser generators, such as 
Boost.Spirit  allow generating output, but require you to write another set of semantic actions.
